---
title: "Setting Up Your System"
author: "Eric C. Anderson"
date: "1/2/2020"
output: 
  slidy_presentation:
    css: csu-con-gen-slides.css
    footer: "&nbsp;&nbsp;&nbsp;Written by Eric C. Anderson using rmarkdown slidy. &nbsp;&nbsp; Click <a href='https://github.com/eriqande/CSU-con-gen-comp-2020/edit/master/docs/slides-and-things/setting-up-your-system.Rmd'>here</a> to suggest edits via GitHub."
    includes:
      before_body: csu-con-gen-doc-prefix.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## R and Rstudio Installations

These are pretty easy to install on either Windows or a Mac.  The directions
are largely the same for either system.


### R

- On December 12, 2019, version 3.6.2 was released.
- That is the version we will be using.

1. Go to [https://cran.r-project.org/](https://cran.r-project.org/) and find the download link for your computer system.
1. Download it and install it. Just do a standard install, with all the default options.

**On Windows:** You should also go back to the cran downloads page and download and install the **Rtools35.exe**.  This will let you compile and build R packages on your computer.


### RStudio

1. We want the latest version of RStudio. 
1. Get the free desktop version from [https://rstudio.com/products/rstudio/download/](https://rstudio.com/products/rstudio/download/) 
1. install the appropriate one for your OS.


### Some R Packages

- **bookdown:** This package is what I used to create these course notes.  Getting it automatically installs a lot of 
other packages that are useful for authoring reproducible research.
    ```{r get-bd, eval=FALSE, echo=TRUE}
    install.packages("bookdown")  
    ```
- **tinytex:** If you don't already have LaTeX installed, do the following so that you can be guided in maintaining a LaTeX distribution.  Basically, with this package installed,
if you try to create PDF output from an RMarkdown document, this package will detect that
and give you a message about how to install a LaTeX system.
    ```{r, get-tt, eval=FALSE, echo=TRUE}
    install.packages("tinytex")  
    ```

- **other R packages** that we will want to have, namely, the package `tidyverse`, which
installs a number of Hadley Wickham's packages, and also a package called `remotes` that
makes it easy to install packages from GitHub and other repositories..
    ```{r, othpack, eval=FALSE, echo=TRUE}
    install.packages(c("tidyverse", "remotes"))  
    ```


## Getting a Bash shell

### On a Mac

- This comes standard with OS X on a Macintosh. 
- You can access it from the Terminal app, and it also works just fine, driving it from RStudio...

### On Windows

Getting Bash on Windows is a little tougher than on a Mac.  I recommend
here that you get the **Git Bash** system, which gives you a serviceable
Unix-like command line for Windows.  And you get Bash in the process as well.  Yay!

You can follow these steps:

1. Download the Git For Windows Installer at [https://git-scm.com/download/win](https://git-scm.com/download/win)
1. Run the installer by double-clicking on the downloaded file and by following the steps bellow:
    - GNU GPL screen. Click on “Run”.
    - Select components.  Click on “Next” (use defaults).
    - Editor selection.  Choose "Use the Nano Editor by default." Click on “Next”.
    - Adjusting your PATH environment. Make sure "Git from the command line and also from third party software." is selected.  Click on “Next”.
    - Choosing HTTPS transport backend.  Stick with the default, Use the OpenSSL library." Click on “Next”.
    - Line endings.  Make sure Checkout Windows-style commit Unix style is selected (which is the default). Click on “Next”.
    - Use Windows' default console window.  Select the second option for Use Windows’ default console window and click on “Next”.
    - Configuring extra options.  Keep it as the default (top top boxes checked). Click on “Next”.
    - Experimental options.  Leave it unchecked. Click on “Install”.
    - When the install is complete, click on “Finish”.
1. Finally, in order to allow R to call system commands via this Git Bash shell (and to
allow shell code blocks in RMarkdown to work properly), you need to add its location
to your Windows Path Environment Variable.  Good directions for how to add a windows
path variable can be found [here](https://helpdeskgeek.com/windows-10/add-windows-path-environment-variable/).
Follow those directions and add the path `C:\Program Files\Git\bin` if it
is not already there.


## Getting git and telling rstudio where to find it

### On Windows

If you successfully installed Git for Windows, then you have git. You just have to
let RStudio where to find it.  

1. Go to Tools -> Global Options.
2. Select Git/SVN.
3. Make sure that the box next to "Enable version control interface for RStudio projects"
is checked.
4. Under Git executable you should see `C:/Program Files/Git/bin/git.exe`. If it is empty, try adding that text.


### On a Mac

You might already have git.  To find out do this:

- If you are using a Mac with a reasonably new OS,
you should be able to just open the Terminal application (`/Applications/Utilities/Terminal`) and type "git"
at the command line.  If you have `git` it will say something that starts like:
    ```{sh, eval=FALSE, echo=TRUE}
    usage: git [--version] [--help] [-C <path>] [-c name=value]
           [--exec-path[=<path>]] [--html-path] [--man-path] [--info-path]
           [-p | --paginate | --no-pager] [--no-replace-objects] [--bare]
           [--git-dir=<path>] [--work-tree=<path>] [--namespace=<name>]
           <command> [<args>]

These are common Git commands used in various situations:

start a working area (see also: git help tutorial)
   clone      Clone a repository into a new directory
   etc. etc. etc.
    ```
If you do not have `git` then it should pop up a little thing asking if you would like to install
a reduced set of developer tools.  You do.  Click OK.  **NOTE** Instead of a pop up it might say something
like, "xcrun Error: invalid active developer path. etc. etc...".  In that case, you can install a fresh set 
of command line tools by typing this at the command line:
    ```{sh, eval=FALSE, echo=TRUE}
    xcode-select --install
    ```

### On Linux

- If you are using Linux then we will assume you know how to get `git` or that you already have it.


## git configurations (name, keychain, editor)

There a few configurations to do for git:

### Name and email

When you use git, it stamps the changes with your name and email.  You need
to let it know those things, if you haven't already configured things.
You need to do this on both a PC and a Mac.

1. In RStudio, choose `Tools -> Terminal -> New Terminal`. That will pop up a Unix shell
in a new window in RStudio's lower left pane.
2. Type the following lines into that terminal window replacing John Doe and his email with your name
and your email, hitting return after each.
    ```{sh, eval=FALSE, echo=TRUE}
    git config --global user.name "John Doe"
    git config --global user.email johndoe@example.com
    ```


### Credential helper

On a Mac, there is one further step that will help you not have to
re-type your password every time git communicates with GitHub.  (On Windows,
if you installed Git/Bash for Windows according to the directions given earlier, 
this type of feature should already be enabled.
    ```{sh, eval=FALSE, echo=TRUE}
    git config --global credential.helper osxkeychain
    ```


## Get an account on GitHub

If you don't already have an account on GitHub, go to [github.com](github.com) and click the "sign up" link near upper right 
of the page.  It is pretty self-explanatory.  Go ahead and get a **free** account.  There is nothing to pay for here!

### Private repositories

_If you are a graduate student_ and you do not feel comfortable posting your data on a public site like
GitHub, then you should request some private repositories from GitHub.  GitHub has a great deal for academic
users like students: free private repositories.  Please go to [https://education.github.com/pack](https://education.github.com/pack)
to sign up for your free student pack.  


## Check out an Rstudio project

I am going to have the students check out a small project and run
a notebook to make sure that everything is working correctly.


